---
title: "Data aggregation and reporting delays"
output: rmarkdown::html_vignette
author: "Chi Zhang"
date: "2022-02-23"
vignette: >
  %\VignetteIndexEntry{Data aggregation and reporting delays}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides information on aggregation death registration data. 

NOTE: the data example here are **simulated**, not real. 

```{r setup}
library(nowcast)
library(magrittr)
```

## Individual level daily mortality reporting

The raw data typically has **DOE** (date of event/death), **DOR** (date of registration). Sometimes it also contains additional information such as location and age.

```{r}
data_fake_nowcasting_raw
```






## Weekly aggregation by location

We need to specify a time point until when we stop counting. This is usually the last or latest date available. 

The time unit here is **week**. The events (deaths) happened in a natural week do not necessarily get counted within the same week. Therefore we create a data table that counts the number of events:

- happened in this week
- happened and were registered within this week (lag 0)
- happened and were registered within this and the next week (lag 1)
- ...

The number are cumulatives.


These two concepts are different: 
- deaths happened and registered within the same week
- deaths registered in this week (mixture of current and delayed registrations from previous weeks)


For convenience, we only count the events on a weekly basis, i.e. if an event happened on a Sunday and is counted on the next Monday, the **delay in week** (within the next week) is used, rather than the delay in day (one day delay). 

In the future version, we might investigate this in detail.



### Without weekly percentile 

```{r}
weekly_counts <- count_weekly_reporting(
  data = data_fake_nowcasting_raw, 
  aggregation_date = as.Date('2020-01-01'), 
  max_week_delay = 4)
weekly_counts
```




### With weekly percentile 

If desired, weekly percentile of the cumulative registrations can be easily computed. This is used for explorative analysis so that we can see how much percent of events are registered completely.

```{r}
weekly_counts_p <- count_weekly_reporting(
  data = data_fake_nowcasting_raw, 
  aggregation_date = as.Date('2020-01-01'), 
  max_week_delay = 4, 
  keep_weekly_prob = T)
weekly_counts_p
```







## Old aggregation data

The old version of weekly aggregation is kept here for reference

```{r}
data_fake_nowcasting_aggregated
```



Another data (nation, simulated) looks like this. It is kept for now because `fit_attrib()` requires it as an example.


```{r}
data_fake_nation
```



