---
title: "Aggregation of mortality data"
output: rmarkdown::html_vignette
author: "Chi Zhang"
date: "2022-02-23"
vignette: >
  %\VignetteIndexEntry{Aggregation of mortality data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides information on death registration data. 

NOTE: the data example here are **simulated**, not real. 

```{r setup}
library(nowcast)
library(magrittr)
```


The raw data typically has DOE (date of event/death), DOR (date of registration). Sometimes it also contains additional information such as location and age.

```{r}
data_fake_nowcasting_raw
```


After aggregation the data looks like this 


```{r}
data_fake_nowcasting_aggregated
```



Another data (nation, simulated) looks like this. It is kept for now because `fit_attrib()` requires it as an example.


```{r}
data_fake_nation
```



